The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-07-17 script/en/chapters/paper_and_pencil.tex
added figure to paper_and_pencil appendix
paper and pencil: converted pseudo appendix in to section; defined sagecode float (en only) for sage code examples

% ---------------------------------------------------------------------------
\newpage
\section*{Appendix A: Examples using Sage}
\addcontentsline{toc}{section}{Appendix A: Examples using Sage}
\index{Sage}

\subsection*{Transposition ciphers using Sage}

Tranposition ciphers are implemented in the class
\begin{center}
\verb!sage.crypto.classical.TranspositionCryptosystem!
\end{center}
To construct and work with a transposition cipher, we first need to
determine the alphabet that contains our plaintext and ciphertext
spaces. Typically, this alphabet will be the upper-case letters of the
English alphabet, which can be accessed via the function
\begin{center}
\verb!sage.monoids.string_monoid.AlphabeticStrings!
\end{center}
We then need to decide on the block length of a block permutation,
which is the length of the row vector to be used in encryption and
decryption. This row vector is our key, and it specifies a permutation
of a plaintext.

@@

\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
sage: # transposition cipher of block length 14
sage: T = TranspositionCryptosystem(AlphabeticStrings(), 14)
sage: # our encryption/decryption key
sage: key = [2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 1]
sage:
sage: # encrypt a plaintext
sage: E = T(key)
sage: C = E(T.encoding("a b c d e f g h i j k l m n")); C
BCDEFGHIJKLMNA
sage:
sage: # decrypt ciphertext
sage: keyInv = [14, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13]
sage: D = T(keyInv)
sage: D(C)
ABCDEFGHIJKLMN
\end{Verbatim}

@@

\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
sage: T = TranspositionCryptosystem(AlphabeticStrings(), 14)
sage: # encrypt by shifting to the left by 2 letters
sage: key = [3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 1, 2]
sage: E = T(key)
sage: C = E(T.encoding("a b c d e f g h i j k l m n")); C
CDEFGHIJKLMNAB
sage:
sage: # decrypt by shifting to the left by 12 letters
sage: keyInv = [13, 14, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12]
sage: D = T(keyInv)
sage: D(C)
ABCDEFGHIJKLMN
\end{Verbatim}

@@

\subsection*{Substitution ciphers}

@@

\noindent The following code sample uses Sage to construct a substitution
cipher with a random key. A random key can be generated using the
method \verb!random_key()! of the class
\verb!SubstitutionCryptosystem!. Different keys determine different
substitution ciphers. So each call to \verb!random_key()! returns
different results. First, we perform the encryption process with a
random key.

@@

\subsection*{Caesar cipher}

@@

\subsection*{Substitution with symbols}

@@

\subsection*{Vigen{\`e}re cipher}

@@

\subsection*{Hill cipher}

>>>>

% ---------------------------------------------------------------------------
\newpage
\section{Examples using Sage}
\index{Sage}

\subsection{Transposition ciphers using Sage}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.5]{figures/encryption-decryption-en}
\caption{XXX} 
\label{XXX}
\end{center}
\end{figure}

Tranposition ciphers are implemented in the class
\begin{center}
\verb!sage.crypto.classical.TranspositionCryptosystem!
\end{center}
To construct and work with a transposition cipher, we first need to
determine the alphabet that contains our plaintext and ciphertext. 
Typically, this alphabet will be the upper-case letters of the
English alphabet, which can be accessed via the function
\begin{center}
\verb!sage.monoids.string_monoid.AlphabeticStrings!
\end{center}
We then need to decide on the block length of a block permutation,
which is the length of the row vector to be used in encryption and
decryption. This row vector is our key, and it specifies a permutation
of a plaintext.

@@

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
sage: # transposition cipher of block length 14
sage: T = TranspositionCryptosystem(AlphabeticStrings(), 14)
sage: # our encryption/decryption key
sage: key = [2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 1]
sage:
sage: # encrypt a plaintext
sage: E = T(key)
sage: C = E(T.encoding("a b c d e f g h i j k l m n")); C
BCDEFGHIJKLMNA
sage:
sage: # decrypt ciphertext
sage: keyInv = [14, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13]
sage: D = T(keyInv)
sage: D(C)
ABCDEFGHIJKLMN
\end{Verbatim}
\caption{Caption 1}
\end{sagecode}

@@

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
sage: T = TranspositionCryptosystem(AlphabeticStrings(), 14)
sage: # encrypt by shifting to the left by 2 letters
sage: key = [3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 1, 2]
sage: E = T(key)
sage: C = E(T.encoding("a b c d e f g h i j k l m n")); C
CDEFGHIJKLMNAB
sage:
sage: # decrypt by shifting to the left by 12 letters
sage: keyInv = [13, 14, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12]
sage: D = T(keyInv)
sage: D(C)
ABCDEFGHIJKLMN
\end{Verbatim}
\caption{Caption 2}
\end{sagecode}

@@

\subsection{Substitution ciphers}

@@

\noindent The following code sample uses Sage to construct a substitution
cipher with a random key. A random key can be generated using the
method \verb!random_key()! of the class
\texttt{SubstitutionCrypto\-system}. Different keys determine different
substitution ciphers. So each call to \verb!random_key()! returns
different results. First, we perform the encryption process with a
random key.

@@

\subsection{Caesar cipher}

@@

\subsection{Substitution with symbols}

@@

\subsection{Vigen{\`e}re cipher}

@@

\subsection{Hill cipher}
